---
title: "Flowers_classification"
author: "Zhaomingli"
date: "26/06/2017"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE,message=TRUE,warning=FALSE,error=FALSE,results=TRUE,eval=FALSE,tidy=TRUE,results='hide'}
knitr::opts_chunk$set(echo = TRUE)
```

## Read and explore dataset 

### Summary of the dataset:
```{r}
flowerDataset <- read.csv("./flowers.csv")
summary(flowerDataset)
```
### Preprocessing dataset

The "root"" feature has 1791 missing values which is not trival relative to the total size of sample, so we must handle these missing values at first.
```{r  message=FALSE,error=FALSE,warning=FALSE}
flowerDataset$root[flowerDataset$root=="?"] <- NA
library(mice)
library(VIM)
mice_plot <- aggr(flowerDataset, col=c('green','yellow'),
                    numbers=TRUE, sortVars=TRUE,
                    labels=names(flowerDataset), cex.axis=.7,
                    gap=3, ylab=c("Missing data","Pattern"))
# imputed data with 50 iterations and predictive mean mathing method.
imputed_Data <- mice(flowerDataset, maxit = 50, method = 'pmm', seed = 500)
completeData <- complete(imputed_Data)
#Check the missing in imputed dataset
summary(completeData)
```

Remove the feature which values is identical among observations
```{r}
flowerDataset$seed.size <- NULL
```
## Train the model
### Split the sample data into training and testing.
```{r  message=FALSE,error=FALSE,warning=FALSE}
library(caret)
library(caTools) 
set.seed(101) 
sample = sample.split(completeData$class, SplitRatio = .7)
completeDataTrain = subset(completeData, sample == TRUE)
completeDataTest = subset(completeData, sample == FALSE)

dataTrainClass <- completeDataTrain$class
dataTestClass <- completeDataTest$class
completeDataTrain$class<-NULL
completeDataTest$class<-NULL

#create cross validation folds 
crossValidationFolds<-createMultiFolds(dataTrainClass,k=5,times=1)
#create fold object
trainControlObject<-trainControl(method="repeatedcv",number=5,repeats=1,index=crossValidationFolds)
```

### Train a random forest modle
```{r  message=FALSE,error=FALSE,warning=FALSE}

rfModel<-train(x=completeDataTrain,y=dataTrainClass,method="rf",trControl=trainControlObject,tuneLength=3)
plot(varImp(rfModel),main="Random Forest - Variable Importance")

rfPredicted<-predict(rfModel,completeDataTest)

rfTestCompare<-data.frame(orignal=dataTestClass,predicted=rfPredicted)

confusionMatrix(table(rfTestCompare$orignal,rfTestCompare$predicted)) 

```


### Train a logistic regression model
```{r message=FALSE,error=FALSE,warning=FALSE}
#Reset trainset
completeDataTrain = subset(completeData, sample == TRUE)
#Using three important features (from RF) 
lrModel <- train(class ~ scent + seed.dispersal + stem.surface.primary + fruit.size + stem.surface.secondary + fruit.color,  data=completeDataTrain, method="glm", family="binomial",
                 trControl = trainControlObject, tuneLength = 5)
predictor = predict(lrModel, newdata=completeDataTest)
confusionMatrix(data=predictor, dataTestClass)
```

## Prediction
```{r  message=FALSE,error=FALSE,warning=FALSE}
toPredictFlowerDatasetOrig <- read.csv("./flowers_classify.csv")
toPredictFlowerDataset <- toPredictFlowerDatasetOrig
toPredictFlowerDataset$root[toPredictFlowerDataset$root=="?"] <- NA
toPredictFlowerDataset$seed.size <- NULL
#Handel missing value
toPredictFlowerDatasetImputed <- mice(toPredictFlowerDataset, maxit = 50, method = 'pmm', seed = 500)
toPredictFlowerDataset <- complete(toPredictFlowerDatasetImputed)
summary(toPredictFlowerDataset)
resultRFPredicted<- predict(lrModel, newdata=toPredictFlowerDataset)
toPredictFlowerDatasetOrig$class <- resultRFPredicted
summary(toPredictFlowerDatasetOrig)
write.csv(toPredictFlowerDatasetOrig,file="./flowers_result.csv",row.names=FALSE)
```
## Conclusion
scent is the most indicative of an poisinous flower. There are 331 edible and 409 poisonous flowers.